---
output:
    html_notebook:
        toc: true
        toc_float: true
editor_options:
  chunk_output_type: inline
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(magrittr)
library(devtools)
library(sessioninfo)
```

# `stats19` - package review

## **Reviewer:** [\@adamhsparks](https://github.com/adamhsparks)

### Review Submitted: 28.12.2018
**`r cat(sprintf("**Last updated:** %s", Sys.Date()))`**

***

<br>

This report contains documents associated with the review of **rOpenSci** submitted package:

### **`stats19`: ropensci/software-review**  issue [https://github.com/ropensci/software-review/issues/266](https://github.com/ropensci/onboarding/issues/266).

<br>

## Package info

**Description:**

Tools to help process and analysie the UK road collision data also known as STATS19.

**Author:** `r c(
    person("Robin", "Lovelace", email = "rob00x@gmail.com", role = c("aut", "cre"),
    comment = c(ORCID = "0000-0001-5679-6536")),
    person("Malcolm", "Morgan", email = "m.morgan1@leeds.ac.uk", role = c("aut")),
    person("Layik", "Hama", email = "layik.hama@gmail.com", role = c("aut"),
    comment = c(ORCID = "0000-0003-1912-4890")),
    person("Mark", "Padgham", role=c("aut"))
  )`

**repo url:** <https://github.com/ITSLeeds/stats19>

**website url:** <>

## Review info


#### See [reviewer guidelines](https://ropensci.github.io/dev_guide/reviewerguide.html) for further information on the rOpenSci review process.

**key review checks:**

- Does the code comply with **general principles in the [Mozilla reviewing guide](https://mozillascience.github.io/codeReview/review.html)**?
- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
- Are there **improvements** that could be made to the **code style?**
- Is there **code duplication** in the package that should be reduced?
- Are there **user interface improvements** that could be made?
- Are there **performance improvements** that could be made?
- Is the [**documentation**](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) **clear and sufficient**?

Please be respectful and kind to the authors in your reviews. The rOpenSci [code of conduct](https://ropensci.github.io/dev_guide/policies.html#code-of-conduct) is mandatory for everyone involved in our review process.

***

### session info


```{r sessionInfo}
session_info()
```


```{r pkg_dir, echo = F}
pkg_dir <- "/Users/adamsparks/Documents/Reviews/stats19"
```

## Test installation

### test local `stats19` install:

```{r test-local}
install(pkg_dir, dependencies = T, build_vignettes = T)
```

```{r github-rm}
remove.packages("stats19")
```
#### **comments:**

Local version installed with no issues.

***

### test install of `stats19` from GitHub with:

```{r test-github}
remotes::install_github("ITSLeeds/stats19", dependencies = T, build_vignettes = T)
```

#### **comments:**

GitHub version installs with no issues.

***



## Check package integrity

### run checks on `stats19` source:

```{r check-checks}
devtools::check(pkg_dir)
```
#### **comments:**

No errors, warnings or notes.

***

### run tests on `stats19` source:

```{r check-tests}
devtools::test(pkg_dir)
```
#### **comments:**

One test fails:

```r
est-dl_stats19.R:29: failure: dl_stats19 works for a type
`dl_stats19(type = "accid")` produced unexpected messages.
Expected match: More than one file found, selecting the first
Actual values:
* Multiple matches. Which do you want to download?
```

```r
test_that("dl_stats19 works for a type", {
  skip_download()
  # already downloaded
  expect_message(dl_stats19(type = "accid"),
                 "More than one file found, selecting the first")
})
```

***


### check `stats19` for goodpractice:

```{r test-goodpractice}
goodpractice::gp(pkg_dir)
```
#### **comments:**

<!-- record comments on goodpractice here -->

***

## Check package metadata files

### inspect

- #### [README](https://github.com/ITSLeeds/stats19)
- #### [DESCRIPTION](https://github.com/ITSLeeds/stats19/blob/master/DESCRIPTION)
- #### [NAMESPACE](https://github.com/ITSLeeds/stats19/blob/master/NAMESPACE)

### spell check

```{r spell-check}
devtools::spell_check(pkg_dir)
```

#### **comments:**

- Spellcheck returns some issues that should be addressed, UK English notwithstanding.

- The Description should include a link to the DfT website.

- Language field should be set, `Language: en-GB`

***

## Check documentation

online documentation: **<>**

* Is the [documentation](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) clear and sufficient?

### test `stats19` function help files:

```{r test-help}
help(package = "stats19")
```

#### **comments:**

- README is clear on installation and how/why it's used.

- The contributing guidelines on GitHub mention a `pkgdown` website, this does not exist

- Use `remotes::install_github()` in place of `devtools::install_github()` in README.

- The code style is inconsistent in the README.Rmd file in the code chunks, e.g. line 85 is missing space around `=`.

- The example in the README showing two steps seems necessarily confusing to new users. If there is a good reason for having the raw data in R, document in a vignette why this is useful and show the two-step process, but if normal users won't do this, I wouldn't show it in the quick-start.

- Line 43 of README uses inconsistent "(" around the phrases with the other two `read_*` function description.

***

### test `stats19` vignettes:

```{r test-vignettes}
vignette(package = "stats19")
```

#### **comments:**

- Run spell-check on it.

- The term "attach"" has a specific meaning in R. Suggest rewording the portion about installation and loading the package to omit the use of "attach", since you're not using `attach()` in the R sense (and really shouldn't use it anyway).

- I would describe why a user might want or need to install the Development version from GitHub in the vignette. Presumably if they are reading the vignette, they've already installed the package from CRAN (in the future).

- Try to consistently use `function()` to identify functions in the vignette text. This also means that if/when you use pkgdown to build a site, the functions are linked to the help file.

- In the introduction, the description of why there are `read_*()` and `format_*()` functions is confusing. To me, it reads as if `format` is only a parameter for `read_*()` in the introduction. I was left wondering why it's documented there or why the `format_*()`s even exist until I reached the end of the vignette.

- There is a comma out of place in Vignette,

>Format: Each of the read_*() functions has a format parameter which, when TRUE, adds

should be 

>Format: Each of the read_*() functions has a format parameter, which, when TRUE, adds 

- I'm unsure about including a package that's not on CRAN in the vignette (`ukboundaries`), something like this should be listed in Suggests, but it's not on CRAN, @skcott do you have any thoughts?

- The first figures in the `sf` section after the join aren't immediately clear to me. The axis lack labels, I'm not really sure what I'm looking at.



***

## Test functionality:

- Are there **user interface improvements** that could be made?
- Are there **performance improvements** that could be made?

```{r free-style}
library("stats19")
```

```{r parse-functions}
exports <- ls("package:stats19")
exports
```

<!-- experiment with package functions -->

```{r exp-chunk}
library(stats19)
x = read_accidents()

```

#### **comments:**

The example for `format_sf()` does not work properly when executed as shown.

***

## Inspect code:

- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
    * good [function & variable naming?](https://ropensci.github.io/dev_guide/building.html#function-and-argument-naming)
    
      - Yes
    
    * good [dependency management](https://ropensci.github.io/dev_guide/building.html#package-dependencies)?
    
      - `roxygen2` is not listed
      - The package does not pass `R CMD check`. *curl*, *readxl* and *tibble* are all listed as an Imports in DESCRIPTION but not imported from. With *curl* being used in tests, this means it should be in Suggests, I think. The others should be removed.
    
- Are there **improvements** that could be made to the [**code style?**](https://ropensci.github.io/dev_guide/building.html#code-style)

  - Whitespaces useage is inconsistent in some spots
  
- Is there **code duplication** in the package that should be reduced?

  - None noted

```{r inspect-code}
pkgreviewr::pkgreview_print_source("stats19")
```



<br>
<br>

#### **comments:**

- Per rOpenSci policy, avoid start-up messages, rely on the documentation for
citation information:
https://ropensci.github.io/dev_guide/building.html#recommended-scaffolding.

- Avoid long lines >80 chars

- Inconsistent use of white spaces in code, see `find_file_name()` `if` statements for examples.

- The package does not pass `R CMD check`. *curl*, *readxl* and *tibble* are all listed as an Imports in DESCRIPTION but not imported from. With *curl* being used in tests, this means it should be in Suggests, I think. The others should be removed.

- I don't think it's good form to have an example that won't work on Windows in the help file for `stats19_schema`, from data.R - line 17? Most of what I see there would be better served in a `data_raw` folder showing how the data were created with the documentation actually documenting what the variables are not how they were created, see <http://r-pkgs.had.co.nz/data.html> and for an example, <https://github.com/ropensci/GSODR/tree/master/data-raw>.

- I would suggest to use proper formatting in help files, when naming packages, e.g. \pkg{stats19} and when referring to documented functions or data, e.g. \code{\link{stats19_schema}}, or with single quotes around abbreviations, e.g. 'DfT'. @ColinFay has an excellent page that outlines the formatting options and when/how to use them, <https://colinfay.me/writing-r-extensions/writing-r-documentation-files.html>. This will greatly enhance the users' experience when using the help files by making them more readable.

- I also would suggest making use of `@seealso` in documentation. For example, the `dl_stats19()` example works great in the help files, but from there I have the data but it's not in R. Using the `@seealso` you can let the user know about the `read_*()` functions.

- I downloaded files using `dl_stats19()`, selecting "Casualties", and then ran `read_accidents()` and got 
```r
Error in check_input_file(filename = filename, type = "accidents", data_dir = data_dir,  : 
  Change data_dir, filename, year or run dl_stats19() first.
```
Is it possible to be more descriptive and say that I've used the wrong `read_*()` based on the file/data found and offer to import it?

- Missing "." after "e.g." in dl.R on lines 8 and 9, there may be others that I didn't spy.

- Capitalisation in help files is inconsistent, e.g. lines 123-125 of read.R, parameter descriptions are mixed upper and lower case for first word after parameter itself. This applies to other functions where the descriptions are given in all lower case for other functions or upper case.

- Testing the functionality, I get this, when I expect it to tell me that `deaths` is not a valid input. But then when I hit escape, I expect it simply exit, not provide a warning message on the way out as well.
```r
dl_stats19(year = 1979, type = "deaths")
No files of that type found for that year.
This will download 240 MB+ (1.8 GB unzipped).
Files identified: Stats19-Data1979-2004.zip

Download now (y = enter, n = esc)? 

Warning message:
In find_file_name(years = year, type = type) :
  Coordinates unreliable in this data.
```

- I got caught out when using the interactive features. I read "y = enter" but hit "y" thinking that would work as well as hitting "enter", but R cancelled the operation anyway just as if I'd hit "esc"

- Per a recent conversation with CRAN, you should use `donttest()` rather than `dontrun()` for examples you don't want to be run on CRAN. Then set .travis.yml to run them by using `r_check_args: --as-cran --run-donttest`.  **This may not be appropriate in all cases, e.g. interactive functions.**

- When validating user inputs and using `stop()` it's nice to use `call. = FALSE` to simplify the error message that the user receives.

- Consider using [`hoardr`](https://ropensci.github.io/hoardr/) for managing user-saved files on disk that aren't in `tempdir()`?

- When using `utils::download.file()`, you should use `mode = "wb"` or Windows users may end up with corrupted downloads in my experience. `curl::curl_download()` does the same thing but uses more updated ways of doing it and defaults to using a binary mode (wb).

- I don't think that there is much need for the `Attempting download from` or `Reading in: ` message. If it takes that long, I would suggest to use a progress bar to show progress. But this is just a personal observation.

- Consider setting up a `pkgdown` site? It's easy to do and you can automate deployment with your Travis-CI so it's less to remember.

## Review test suite:

See guidance on [testing](https://ropensci.github.io/dev_guide/building.html#testing) for further details.

### test coverage

```{r pkg_coverage}
covr::package_coverage(pkg_dir)
```

### inspect [tests](https://github.com/ITSLeeds/stats19/blob/master/tests/testthat)

#### **comments:**

- I'm unclear how the interactive portion of the package functions is handled in testing? There are ways to handle this, but I don't see any implemented and when I run `devtools::test()` I'm asked to provide my own input.

- Suggest using `skip_on_cran()` since some of the tests can take some time to execute due to download times.

***
