---
output:
    html_notebook:
        toc: true
        toc_float: true
editor_options:
  chunk_output_type: inline
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(magrittr)
```

# `stats19` - package review

## **Reviewer:** [\@adamhsparks](https://github.com/adamhsparks)

### Review Submitted:
**`r cat(sprintf("**Last updated:** %s", Sys.Date()))`**

***

<br>

This report contains documents associated with the review of **rOpenSci** submitted package:

### **`stats19`: ropensci/software-review**  issue [\#NA](https://github.com/ropensci/onboarding/issues/).

<br>

## Package info

**Description:**

Tools to help process and analysie the UK road collision data also known as STATS19.

**Author:** `r c(
    person("Robin", "Lovelace", email = "rob00x@gmail.com", role = c("aut", "cre"),
    comment = c(ORCID = "0000-0001-5679-6536")),
    person("Malcolm", "Morgan", email = "m.morgan1@leeds.ac.uk", role = c("aut")),
    person("Layik", "Hama", email = "layik.hama@gmail.com", role = c("aut"),
    comment = c(ORCID = "0000-0003-1912-4890")),
    person("Mark", "Padgham", role=c("aut"))
  )`

**repo url:** <https://github.com/ITSLeeds/stats19>

**website url:** <>

## Review info


#### See [reviewer guidelines](https://ropensci.github.io/dev_guide/reviewerguide.html) for further information on the rOpenSci review process.

**key review checks:**

- Does the code comply with **general principles in the [Mozilla reviewing guide](https://mozillascience.github.io/codeReview/review.html)**?
- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
- Are there **improvements** that could be made to the **code style?**
- Is there **code duplication** in the package that should be reduced?
- Are there **user interface improvements** that could be made?
- Are there **performance improvements** that could be made?
- Is the [**documentation**](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) **clear and sufficient**?

Please be respectful and kind to the authors in your reviews. The rOpenSci [code of conduct](https://ropensci.github.io/dev_guide/policies.html#code-of-conduct) is mandatory for everyone involved in our review process.

***

### session info


```{r sessionInfo}
sessionInfo()
```


```{r pkg_dir, echo = F}
pkg_dir <- "/Users/adamsparks/Documents/Reviews/stats19"
```

## Test installation

### test local `stats19` install:

```{r test-local}
install(pkg_dir, dependencies = T, build_vignettes = T)
```

```{r github-rm}
remove.packages("stats19")
```
#### **comments:**

<!-- record comments on local install here -->

***

### test install of `stats19` from GitHub with:

```{r test-github}
devtools::install_github("ITSLeeds/stats19", dependencies = T, build_vignettes = T)
```

#### **comments:**

<!-- record comments on github install here -->

***



## Check package integrity

### run checks on `stats19` source:

```{r check-checks}
devtools::check(pkg_dir)
```
#### **comments:**

<!-- record comments on checks here -->

***

### run tests on `stats19` source:

```{r check-tests}
devtools::test(pkg_dir)
```
#### **comments:**

<!-- record comments on tests here -->

***


### check `stats19` for goodpractice:

```{r test-goodpractice}
goodpractice::gp(pkg_dir)
```
#### **comments:**

<!-- record comments on goodpractice here -->

***

## Check package metadata files

### inspect

- #### [README](https://github.com/ITSLeeds/stats19)
- #### [DESCRIPTION](https://github.com/ITSLeeds/stats19/blob/master/DESCRIPTION)
- #### [NAMESPACE](https://github.com/ITSLeeds/stats19/blob/master/NAMESPACE)

### spell check

```{r spell-check}
devtools::spell_check(pkg_dir)
```

#### **comments:**

<!-- record comments on metadata files here -->

***

## Check documentation

online documentation: **<>**

* Is the [documentation](https://ropensci.github.io/dev_guide/building.html#documentation) (installation instructions/vignettes/examples/demos) clear and sufficient?

### test `stats19` function help files:

```{r test-help}
help(package = "stats19")
```

#### **comments:**

<!-- record comments on help files here -->

***

### test `stats19` vignettes:

```{r test-vignettes}
vignette(package = "stats19")
```

#### **comments:**

<!-- record comments on vignettes here -->

***

## Test functionality:

- Are there **user interface improvements** that could be made?
- Are there **performance improvements** that could be made?

```{r free-style}
library("stats19")
```

```{r parse-functions}
exports <-ls("package:stats19")
exports
```

<!-- experiment with package functions -->

```{r exp-chunk}


```

#### **comments:**

<!-- record comments on package experimentation here -->

***

## Inspect code:

- Does the package **comply with the [ROpenSci packaging guide](https://ropensci.github.io/dev_guide/building.html)**?
    * good [function & variable naming?](https://ropensci.github.io/dev_guide/building.html#function-and-argument-naming)
    * good [dependency management](https://ropensci.github.io/dev_guide/building.html#package-dependencies)?
- Are there **improvements** that could be made to the [**code style?**](https://ropensci.github.io/dev_guide/building.html#code-style)
- Is there **code duplication** in the package that should be reduced?

```{r inspect-code}
pkgreviewr::pkgreview_print_source("stats19")
```
**\* might not be suitable for large packages with many exported functions**


<br>
<br>

#### **comments:**

<!-- record comments on package source code here -->


## Review test suite:

See guidance on [testing](https://ropensci.github.io/dev_guide/building.html#testing) for further details.

### test coverage

```{r pkg_coverage}
covr::package_coverage(pkg_dir)

```

### inspect [tests](https://github.com/ITSLeeds/stats19/blob/master/tests/testthat)


#### **comments:**

<!-- record comments on testing suite here -->


***
